---
title: "Continuous-Time Movement Modeling "
author: "Frank Short, Victoria Zdanowicz, Zoe Albert, & Natalia Kelley"
date: "12/10/2021"
output: 
  html_document:
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
---

1. Introduction
      - Random Walk Models
      - Correlated Random Walk Models
      - Continuous-Time Stochastic Process (CTSP) Models 

2. Objectives
      - Explain statistical concept briefly

3. Coding! With explanations
      - Install Packages
      - Load in Movebank GPS data
      - Remove outliers 
      - Produce Variograms for Individuals
          - Assess variograms
      - Produce guess and fit for individuals:
          - Compare fit to variogram
      - Calculate Autocorrelated Kernel Density Estimates (AKDE) 
      - Calculate Overlap (Gaussian and AKDE) of each individual. 

4. Challenges
      - Sample problems with Movebank data

5. Literature Cited

```{r - packages}
library(curl)
library(ctmm)
library(rgdal)
library(rgeos)
library(raster)
library(dplyr)
```


```{r - loading in data}
f <- curl("https://raw.githubusercontent.com/vzdanowicz/AN588_ctmm/main/Site%20fidelity%20in%20cougars%20and%20coyotes%2C%20Utah_Idaho%20USA%20(data%20from%20Mahoney%20et%20al.%202016)-gps.csv")
data <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(data)
```

```{r - ctmm}
couyotes_ctmm <- as.telemetry(data)

```



